---
title: "STAT 425 Case Study R Code"
author: "Wenxiao Yang (wenxiao5), Giang Le (gianghl2), Yuan Chang (changy4)"
output:
  html_document: default
  pdf_document: default
---

<!-- These are the categories to be graded: Intro, regression analysis process, 
diagnotics, remedial measures, correctness of analysis, conclusion, presentation
and organization, tables plots, R code

Yuan: writing: intro, background of data, summary statistics
Giang: writing: diagnostics, model selection
Yuan: writing: conclusion
Wenxiao: coding -->

## Data Overview
```{r}
library(dplyr)
library(faraway)
library(lmtest)
library(MASS)
```

```{r}
# Read in the data
redwines <- read.csv("redwines.csv")
head(redwines)
```

## Full model and Collinearity Check

```{r}
pairs(redwines[,-c(11,12)])
```
```{r}
full.model <- lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide+ total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
summary(full.model)
```

```{r}
round(cor(redwines[,-c(11,12)]),dig=2)
```
We observe $fixed.acidity$ is mildly correlated with $citric.acid$, $density$ and $pH$; $citric.acid$ is mildly correlated with $fixed.acidity$, $volatile.acidity$ and $pH$; $free.sulfur.dioxide$ is mildly correlated with $total.sulfur.dioxide$.
We consider removing $fixed.acidity$, $citric.acid$ and $free.sulfur.dioxide$.
Let's look at $free.sulfur.dioxide$ firstly,
```{r}
model1 <- lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
summary(model1)
```
Great, $total.sulfur.dioxide$'s p-vlaue decreases significantly.
```{r}
anova(model1, full.model)
```
Since F is large, we can drop the $free.sulfur.dioxide$.
```{r}
x = model.matrix(model1)[,-1]
dim(x)
x = x-matrix(apply(x,2,mean),1599,9, byrow=TRUE)
x = x/matrix(apply(x,2,sd),1599,9, byrow=TRUE)
apply(x,2,mean)
apply(x,2,sd)
```
```{r}
e = eigen(t(x) %*% x)
sqrt(e$val[1]/e$val[9])
```
condition number is 5.26 lower than 30, there is no collinearity now.

Then we will use model1 in the following.

## Detect Unusual Observations

### Check High Leverage Points
```{r}
n=dim(redwines)[1]
p=10
lev=influence(model1)$hat
highlev=lev[lev>2*p/n] #find high leverage points
highlev
```

```{r}
halfnorm(lev,6, labs=as.character(1:length(highlev)), ylab="Leverages")
```

### Check Outliers
```{r}
StuR=rstudent(model1)
qt(0.05/(2*n),n-p-1)
```
```{r}
sort(abs(StuR),decreasing=TRUE)[1:10]
```

No outliers

### Check High Influential Points
```{r}
cook=cooks.distance(model1)
max(cook)
```
```{r}
halfnorm(cook,6,labs=as.character(1:length(cook)),ylab="Cook's distances")
```

We have some high leverage points, but none of them are high influential.
We don't have outliers.

## Check Assumptions
### Check Homoscedasticity
```{r}
plot(model1,which=1)
```
```{r}
bptest(model1)
```

Homoscedasticity doesn't hold

### Check Normality
```{r}
plot(model1,which=2)
```
```{r}
hist(model1$residuals)
```
```{r}
ks.test(residuals(model1), y=pnorm)
```
Normality doesn't hold.

## Use Box Cox Transformation
```{r}
min(redwines$alcohol)
```
Make sure all alchohol values are positive.
```{r}
model.transformation=boxcox(model1,lambda=seq(-2, -1, by=0.01))
```

```{r}
model.transformation$x[model.transformation$y==max(model.transformation$y)]
```
$\lambda$ is -1.53
```{r}
model_bx=lm((alcohol^(-1.53)-1)/(-1.53) ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
summary(model_bx)
```
All predictors are significant. We don't need to move any predictors.
Now let's check unusual observations and assumptions again.

## Check Unusual Observations
### Check High Leverages
```{r}
lev_new=influence(model_bx)$hat
highlev_new=lev_new[lev_new>2*p/n] #find high leverage points
highlev_new
```
```{r}
halfnorm(lev_new,6, labs=as.character(1:length(highlev_new)), ylab="Leverages")
```

### Check Outliers
```{r}
StuR_new=rstudent(model_bx)
qt(0.05/(2*n),n-p-1)
```
```{r}
sort(abs(StuR_new),decreasing=TRUE)[1:10]
```
No outliers

### Check High Influential Points
```{r}
cook_new=cooks.distance(model_bx)
max(cook_new)
```
```{r}
halfnorm(cook_new,6,labs=as.character(1:length(cook_new)),ylab="Cook's distances")
```

We have some high leverage points, but none of them are high influential.
We don't have outliers.

## Check Assumptions
### Check Homoscedasticity
```{r}
plot(model_bx,which=1)
```
```{r}
bptest(model_bx)
```

Homoscedasticity doesn't hold

### Check Normality
```{r}
plot(model_bx,which=2)
```
```{r}
hist(model_bx$residuals)
```
```{r}
ks.test(residuals(model_bx), y=pnorm)
```

Normality doesn't hold.
Box-Cox Transformation can't fix the violation to assumptions.

### Check Non-linearity
#### Added-variables plots
```{r}
#fixed.acidity
modela1 = lm(alcohol ~ volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
modela2 = lm(fixed.acidity ~ volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
reg.a = lm(modela1$residuals ~ modela2$residuals)
#volatile.acidity
modelb1 = lm(alcohol ~ fixed.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
modelb2 = lm(volatile.acidity ~ fixed.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
reg.b = lm(modelb1$residuals ~ modelb2$residuals)
#citric.acid
modelc1 = lm(alcohol ~ fixed.acidity + volatile.acidity + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
modelc2 = lm(citric.acid ~ fixed.acidity + volatile.acidity + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
reg.c = lm(modelc1$residuals ~ modelc2$residuals)
#residual.sugar
modeld1 = lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
modeld2 = lm(residual.sugar ~ fixed.acidity + volatile.acidity + citric.acid + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
reg.d = lm(modeld1$residuals ~ modeld2$residuals)
#chlorides
modele1 = lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
modele2 = lm(chlorides ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
reg.e = lm(modele1$residuals ~ modele2$residuals)
#total.sulfur.dioxide
modelf1 = lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + density + pH + sulphates, data=redwines[,-c(12)])
modelf2 = lm(total.sulfur.dioxide ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  density + pH + sulphates, data=redwines[,-c(12)])
reg.f = lm(modelf1$residuals ~ modelf2$residuals)
#density
modelg1 = lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + pH + sulphates, data=redwines[,-c(12)])
modelg2 = lm(density ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + pH + sulphates, data=redwines[,-c(12)])
reg.g = lm(modelg1$residuals ~ modelg2$residuals)
#pH
modelh1 = lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + sulphates, data=redwines[,-c(12)])
modelh2 = lm(pH ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + sulphates, data=redwines[,-c(12)])
reg.h = lm(modelh1$residuals ~ modelh2$residuals)
#sulphates
modeli1 = lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH, data=redwines[,-c(12)])
modeli2 = lm(sulphates ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides +  total.sulfur.dioxide + density + pH, data=redwines[,-c(12)])
reg.i = lm(modeli1$residuals ~ modeli2$residuals)
```

```{r}
plot(modela2$residuals, modela1$residuals)
abline(reg.a, col="red")
```
```{r}
plot(modelb2$residuals, modelb1$residuals)
abline(reg.b, col="red")
```
```{r}
plot(modelc2$residuals, modelc1$residuals)
abline(reg.c, col="red")
```

```{r}
plot(modeld2$residuals, modeld1$residuals)
abline(reg.d, col="red")
```
```{r}
plot(modele2$residuals, modele1$residuals)
abline(reg.e, col="red")
```
```{r}
plot(modelf2$residuals, modelf1$residuals)
abline(reg.f, col="red")
```
```{r}
plot(modelg2$residuals, modelg1$residuals)
abline(reg.g, col="red")
```
```{r}
plot(modelh2$residuals, modelh1$residuals)
abline(reg.h, col="red")
```
```{r}
plot(modeli2$residuals, modeli1$residuals)
abline(reg.i, col="red")
```

All plots look fine except the $residual.sugar$, we try to transform the $residual.sugar$.
```{r}
#residual.sugar
modeld1 = lm(log(alcohol) ~ fixed.acidity + volatile.acidity + citric.acid + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
modeld2 = lm(log(residual.sugar) ~ fixed.acidity + volatile.acidity + citric.acid + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
reg.d = lm(modeld1$residuals ~ modeld2$residuals)
plot(modeld2$residuals, modeld1$residuals)
abline(reg.d, col="red")
```

We can find that $\log({\text{residual.sugar}})$ works much better. Then, Let's try the new model!

## Log-Transformed Model
```{r}
modelN=lm(log(alcohol) ~ fixed.acidity + volatile.acidity + citric.acid + log(residual.sugar) + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
summary(modelN)
```

$volatile.acidity$' p-value increased and is significant at 0.05 significance level but not
at lower levels. Let's try to remove it!
```{r}
modelN2=lm(log(alcohol) ~ fixed.acidity + citric.acid + log(residual.sugar) + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
summary(modelN2)
```

Now let's check unusual observations and assumptions again.

## Check Unusual Observations
### Check High Leverages
```{r}
p=9
levN=influence(modelN2)$hat
highlevN=levN[levN>2*p/n] #find high leverage points
highlevN
```
```{r}
halfnorm(levN,6, labs=as.character(1:length(highlevN)), ylab="Leverages")
```

### Check Outliers
```{r}
StuRN=rstudent(modelN2)
qt(0.05/(2*n),n-p-1)
```
```{r}
sort(abs(StuRN),decreasing=TRUE)[1:10]
```
#652 is an outlier.

### Check High Influential Points
```{r}
cookN=cooks.distance(modelN2)
max(cookN)
```
```{r}
halfnorm(cookN,6,labs=as.character(1:length(cookN)),ylab="Cook's distances")
```
We have some high leverage points and one outlier, but none of them are high influential.

## Check Assumptions
### Check Homoscedasticity
```{r}
plot(modelN2,which=1)
```
```{r}
bptest(modelN2)
```
Homoscedasticity doesn't hold

### Check Normality
```{r}
plot(modelN2,which=2)
```
```{r}
hist(modelN2$residuals)
```
```{r}
ks.test(residuals(modelN2), y=pnorm)
```
Normality doesn't hold.
Still can't fix the violation to assumptions.


We will use Generalized Least Squares next.


## Generalized Least Squares
```{r}
lm.resid=lm(abs(modelN2$residuals)~ fixed.acidity + citric.acid + log(residual.sugar) + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines[,-c(12)])
summary(lm.resid)
```
```{r}
redwines$weight = 1/lm.resid$fitted.values^2
head(redwines)
```
```{r}
ModelGLS = lm(log(alcohol) ~ fixed.acidity + citric.acid + log(residual.sugar) + chlorides +  total.sulfur.dioxide + density + pH + sulphates, data=redwines, weights=weight)
summary(ModelGLS)
```

## Detect Unusual Observations
### Check High Leverage Points
```{r}
p=9
lev=influence(ModelGLS)$hat
highlev=lev[lev>2*p/n] #find high leverage points
highlev
```

```{r}
halfnorm(lev,6, labs=as.character(1:length(highlev)), ylab="Leverages")
```

### Check Outliers
```{r}
StuR=rstudent(ModelGLS)
qt(0.05/(2*n),n-p-1)
```
```{r}
sort(abs(StuR),decreasing=TRUE)[1:10]
```
No ouliers

### Check High Influential Points
```{r}
cook=cooks.distance(ModelGLS)
sort(abs(cook),decreasing=TRUE)[1:10]
```
```{r}
halfnorm(cook,6,labs=as.character(1:length(cook)),ylab="Cook's distances")
```

We have some high leverage points, but none of them are high influential.
No outliers.

## Check Assumptions
### Check Homoscedasticity
```{r}
plot(ModelGLS,which=1)
```
```{r}
bptest(ModelGLS)
```

Homoscedasticity doesn't hold

### Check Normality
```{r}
plot(ModelGLS,which=2)
```
```{r}
hist(ModelGLS$residuals)
```
```{r}
ks.test(residuals(ModelGLS), y=pnorm)
```

Still can't fix the violation to assumptions.















