---
title: "HW9"
author: "Wenxiao Yang"
date: "11/15/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
# 2
## (a)
```{r}
library(faraway)
data(butterfat)
head(butterfat)
```
The factor effects model is as follows:
$$Y_{ijk}= \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij}+\varepsilon_{ijk} $$
$Y_{ijk}$: butterfat for age $i$, breed $j$, cow $k$ 
$\mu$ : mean butterfat for all samples
$\alpha_{i}:$ effect of age $i$ on butterfat
$\beta_{j}:$ effect of breed $j$ on butterfat
$(\alpha \beta)_{i j}$ : interaction term
The error terms satisfy the usual assumption $\varepsilon_{i j k} \sim \mathcal{N}\left(0, \sigma^{2}\right) .$
Sum Constraints: $\sum_{i} \alpha_{i}=0, \sum_{j} \beta_{j} 0, \sum_{i}(\alpha \beta)_{i j}=\sum_{j}(\alpha \beta)_{i j}=0$

## (b)
```{r}
interaction.plot(butterfat$Age, butterfat$Breed, butterfat$Butterfat)
interaction.plot(butterfat$Breed, butterfat$Age, butterfat$Butterfat)
```
Some of the lines are intersect, so interactions are present.

## (c)
```{r}
model1=lm(log(Butterfat)~Age*Breed,data=butterfat)
anova(model1)
```
The hypothesis we test is
$$
\left\{\begin{array}{l}
H_{0}:(\alpha \beta)_{i j}=0, \forall i, j \\
H_{\alpha}: \text { not all }(\alpha \beta)_{i j}=0
\end{array}\right.
$$
Since the $p$-value is large, we conclude that the interaction term is not statistically significant. So, we can remove it from the model.

## (d)
```{r}
model2=lm(log(Butterfat)~Age+Breed,data=butterfat)
```
$$
\left\{\begin{array}{l}
H_{0}:Y_{ijk}= \mu  + \beta_ j+\varepsilon_{ijk} \\
H_{\alpha}: Y_{ijk}= \mu + \alpha_ i + \beta_ j+\varepsilon_{ijk}
\end{array}\right.
$$
```{r}
model3=lm(log(Butterfat)~Breed,data=butterfat)
anova(model3,model2)
```
The p-value is larger than 0.05, then we can't reject null hypothesis. So, we conclude that Age is statistically insigniﬁcant.

$$
\left\{\begin{array}{l}
H_{0}:Y_{ijk}= \mu + \alpha_ i +\varepsilon_{ijk} \\
H_{\alpha}: Y_{ijk}= \mu + \alpha_ i + \beta_ j+\varepsilon_{ijk}
\end{array}\right.
$$
```{r}
model4=lm(log(Butterfat)~Age,data=butterfat)
anova(model4,model2)
```
The p-value is smaller than 0.05, then we reject null hypothesis. So, we conclude that Breed is statistically signiﬁcant.

## (e)
```{r}
anova(model2)
```
$$D=\mu_{1.}-\mu_{2.}$$
```{r}
butterfat['logbutfat']=log(butterfat$Butterfat)
mean.mature=mean(butterfat$logbutfat[butterfat$Age=="Mature"])
mean.2year=mean(butterfat$logbutfat[butterfat$Age=="2year"])
mean.mature-mean.2year
```

$$\hat{D}=\hat{Y}_{1..}-\hat{Y}_{2..}=0.02338223$$
```{r}
0.00745/25
```

$$MSE=0.00745, s^2_{\hat{D}}=\frac{2MSE}{10*5}=0.000298$$
$$D\in0.02338223\pm T_{94}(0.05/2)\sqrt{0.000298}$$

## (f)
```{r}
anova(model3)
```
```{r}
summary(butterfat$Breed)
```
$$
L=\frac{\mu_{1..}+\mu_{2..}}{2}-\frac{\mu_{3..}+\mu_{5..}}{2}
$$
```{r}
mean.1=mean(butterfat$logbutfat[butterfat$Breed=="Ayrshire"])
mean.2=mean(butterfat$logbutfat[butterfat$Breed=="Canadian"])
mean.3=mean(butterfat$logbutfat[butterfat$Breed=="Guernsey"])
mean.5=mean(butterfat$logbutfat[butterfat$Breed=="Jersey"])
(mean.1+mean.2-mean.3-mean.5)/2
```

$$\hat{L}=\frac{1}{2} \bar{Y}_{1 . .}+\frac{1}{2} \bar{Y}_{2 . .}-\frac{1}{2} \bar{Y}_{3 . .}-\frac{1}{2} \bar{Y}_{5 \cdots}=-0.1843879$$
$$s_{\hat{L}}^{2}=\frac{0.00752}{10*2}\left((1 / 2)^{2}+(1 / 2)^{2}+(1 / 2)^{2}+(1 / 2)^{2}\right)=0.000376$$
So, the interval is
$$
L \in(-0.1843879 \pm T_{95}(0.05 / 2) \sqrt{0.000376})
$$



