---
title: "Detecting Unusual Observations"
author: "Alexandra Chronopoulou"
date: "Lecture #8 Example"
output: html_document
---



### Re-visit the Bike Shares Data to check for Unusual Observations

```{r , results='hide'}
library(faraway)
bikeshares <-  read.csv("BikeShares.csv", header=TRUE)
dim(bikeshares)
head(bikeshares)
# We remove columns 1,  7, 8, 9, 10:
bikeshares.reg = bikeshares[,c(-1,-7,-8,-9,-10)]
head(bikeshares.reg)
```

<br>
<hr>
<br>


#### (a) Check for high-leverage points

<br>

```{r}
n=dim(bikeshares.reg)[1]; # sample size
p=4; # 3 predictors we have in the model plus the intercept

bikeshare.mlr = lm(cnt ~ t1 +  hum + wind_speed, data=bikeshares.reg )
# Compute Leverages
lev=influence(bikeshare.mlr)$hat
# Determine which exceed the 2p/n threshold
newlev = lev[lev>2*p/n]

# Prepare a half-normal plot 
halfnorm(newlev, 6, labs=as.character(1:length(newlev)), ylab="Leverages")
```

Which bike shares have high-leverage?


```{r, results=FALSE}
bikeshares.reg[lev > 2*p/n,]
```

<br><br>

#### (b) Check for outliers

<br>

```{r}
# Compute Studentized Residuals
jack=rstudent(bikeshare.mlr);

# The critical value WITH Bonferroni correction is
qt(.05/(2*n), n-p-1) 
# The critical value WITHOUT Bonferroni correction is
qt(.05/2, n-p-1)

# Sort the residuals indescending order to find outliers (if any)
sort(abs(jack), decreasing=TRUE)[1:10]
```

As we can see here, we have 8 outliers, i.e. the values that are higher (in absolute value) of the critical T distribution value with Bonferroni correction/. These are  observations: #4462, #5130, #5139, #4471, #15888, #5140, #15217, #15385.

<br><br>

#### (c) Check for high influential points

<br>

```{r}
# Compute Cook's Distance
cook = cooks.distance(bikeshare.mlr)
# Extract max Cook's Distance
max(cook)

# Prepare a Half Normal Plot of Cook's distances
halfnorm(cook, 6, labs=as.character(1:length(cook)), ylab="Cook's distances")
```


Although there are no high influential points based on the rule-of-thumb, the Cook’s distance for observation #4471 is much larger than the other samples. 

