---
title: "Least Squares Coefficients: Confidence Intervals & Confidence Regions"
author: "Alexandra Chronopoulou"
date: "Lecture 7 Examples"
output: html_document
---


```{r , results='hide'}
bikeshares <-  read.csv("BikeShares.csv", header=TRUE)
dim(bikeshares)
head(bikeshares)
# We remove columns 1,  7, 8, 9, 10:
bikeshares.reg = bikeshares[,c(-1,-7,-8,-9,-10)]
head(bikeshares.reg)
```

<br>
<hr>
<br>





#### Confidence/Prediction Interval, Confidence Region

Use the command **confint** to obtain confidence intervals for regression coefficients.

```{r conf}
bikeshare.mlr = lm(cnt ~ t1 +  hum + wind_speed, data=bikeshares.reg )
confint(bikeshare.mlr)
confint(bikeshare.mlr, 't1', level=0.99)
```

Use the command **ellipse** from the ellipse package to obtain confidence regions. Here we plot the confidence region for (`t1`, `hum`), which is an ellipsoid centered at the LS estimate of their coefficients.

Recall that we tested whether their LS coefficients are zero using the $F$-test earlier. Now, using the duality of Confidence region/interval and hypothesis test, we reject the hypothesis based on whether the point (0,0) is inside the ellipsoid or not.

```{r ellipse, warning=FALSE}
library(ellipse)
library(ggplot2)

CR95 = ellipse(bikeshare.mlr, c(2,3))
CR99 = ellipse(bikeshare.mlr, c(2,3), level=0.99)
CR998 = ellipse(bikeshare.mlr, c(2,3), level=0.998)
dim(CR95)
head(CR95)
myCR = rbind(CR95, CR99, CR998);
myCR = data.frame(myCR); 
myCR
names(myCR) = c("t1","hum"); 
myCR[, 'level']=as.factor(c(rep(0.95, dim(CR95)[1]), 
                              rep(0.99, dim(CR99)[1]), 
                              rep(0.998, dim(CR998)[1])));


ggplot(data=myCR, aes(x=t1, y=hum, colour=level)) + 
  geom_path(aes(linetype=level), size=1.5) + 
  geom_point(x=coef(bikeshare.mlr)[2], y=coef(bikeshare.mlr)[3], shape=3, size=3, colour='red') + 
  geom_point(x=0, y=0, shape=1, size=3, colour='red') 

```

Let’s find the 95 percent confidence interval (CI) and prediction interval (PI) at the mean value of the predictors.


```{r pred1}
# create a data frame on which you would like to predict
bikeshares.reg[,2:5]
meanvalue=apply(bikeshares.reg[,2:5],2,mean)
apply(bikeshares.reg,2,mean)
meanvalue
x=data.frame(t(meanvalue))
predict.lm(bikeshare.mlr,x,interval="confidence",level=0.95)
predict.lm(bikeshare.mlr,x,interval="prediction")
```

<br>

CIs or PIs get wider as we move away from the training data. 

Next, we display the 95% CIs/PIs for the full model when we vary `wind_speed`, while holding the other predictors fixed at the meanvalue.

```{r pred 2}
# First create the data frame on which you need to predict
#range(bikeshares.reg$wind_speed)
#[1]  0.0 56.5
x=matrix(meanvalue[1:3], 57, 3, byrow=TRUE)
colnames(x)=c("t1", "t2", "hum")
x=data.frame(x)

x[, 'wind_speed']=0:56; 
# head(x)

myCI=predict.lm(bikeshare.mlr,x,interval="confidence", level=0.90);
myPI=predict.lm(bikeshare.mlr,x,interval="prediction", level=0.90);


# Form the data frame for plotting
ggplot(data=NULL, aes(x=0:56)) + 
  geom_line(aes(y=myCI[,1], colour="LSfit"), size=1) + 
  geom_line(aes(y=myCI[,2], colour="90% CI"), size=1) +
  geom_line(aes(y=myCI[,3], colour="90% CI"), size=1) +
  geom_line(aes(y=myPI[,2], colour="90% PI"), size=1, linetype=2) +
  geom_line(aes(y=myPI[,3], colour="90% PI"), size=1, linetype=2) + 
  scale_colour_manual("", values=c("LSfit" = "black",
                               "90% CI" = "blue", 
                               "90% PI"="red")) + 
  xlab("wind_speed") +ylab("bike shares") + 
  geom_vline(xintercept = mean(bikeshares.reg$wind_speed), colour="purple", size=1, linetype=3)

```


