---
title: "Final Project R Code"
author: "Wenxiao Yang"
date: "12/7/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
# R Code
```{r}
bubble.data=read.csv("bubblewrap.csv",header=TRUE)
new.data=transform(bubble.data,line_speed=as.factor(line_speed),loading=as.factor(loading))
new.data
```

```{r}
par(mfrow=c(1,2))
plot(rate~line_speed, new.data)
plot(rate~loading, new.data)
```
The factor effects model is as follows:
$$Y_{ijk}= \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij}+\varepsilon_{ijk} $$

```{r}
par(mfrow=c(1,2))
with(new.data, interaction.plot(line_speed,loading,rate))
with(new.data, interaction.plot(loading,line_speed,rate))
```
Interactions are present.

## F-test
```{r}
model1=lm(rate~line_speed*loading,new.data)
summary(model1)
anova(model1)
```

Since the $p$-values are large, we conclude that the interaction terms are not statistically significant. So, we can remove it from the model.

```{r}
model2=lm(rate~line_speed+loading,new.data)
summary(model2)
anova(model2)
```


## Check Assumptions for Model 2
```{r}
par(mfrow=c(1,2))
plot(model2,which=1:2)
```
### Check Homoscedasticity of Model 2
```{r}
library(lmtest)
bptest(model2)
```
Homoscedasticity holds.

### Check Normality of Model 2
```{r}
ks.test(residuals(model2), y=pnorm)
```
Normality doesn’t hold. Try Box-cox transformation.

## Box-cox transformation
```{r}
library(MASS)
```

```{r}
min(new.data$rate)
```
```{r}
model.transformation=boxcox(model2,lambda=seq(3, 6, by=0.1))
```
```{r}
model.transformation$x[model.transformation$y==max(model.transformation$y)]
```
```{r}
par(mfrow=c(1,2))
with(new.data, interaction.plot(line_speed,loading,(rate^5.5-1)/5.5))
with(new.data, interaction.plot(loading,line_speed,(rate^5.5-1)/5.5))
```
```{r}
model.bc=lm((rate^5.5-1)/5.5~line_speed*loading,new.data)
summary(model.bc)
anova(model.bc)
```
We can find the interaction term is still insignificant.
```{r}
model3=lm((rate^5.5-1)/5.5~line_speed+loading,new.data)
summary(model3)
anova(model3)
```
## Check Assumptions for Model 3
```{r}
par(mfrow=c(1,2))
plot(model3,which=1:2)
```
### Check Homoscedasticity of Model 3
```{r}
bptest(model3)
```
Homoscedasticity holds.

### Check Normality of Model 3
```{r}
ks.test(residuals(model3), y=pnorm)
```
The Box-Cox transformation can't fix Normality problem. However, the model3's $R^2$ is higher, so we use model3.


## Tukey’s Paired Comparison
```{r}
TukeyHSD(aov((rate^5.5-1)/5.5~line_speed+loading,new.data), "line_speed")
```

```{r}
myCIs.line_speed = TukeyHSD(aov((rate^5.5-1)/5.5~line_speed+loading,new.data), "line_speed")
plot(myCIs.line_speed)
```
```{r}
TukeyHSD(aov((rate^5.5-1)/5.5~line_speed+loading,new.data), "loading")
```

```{r}
myCIs.loading = TukeyHSD(aov((rate^5.5-1)/5.5~line_speed+loading,new.data), "loading")
plot(myCIs.loading)
```
