---
title: "Partial F Tests & Permutation Tests"
author: "Alexandra Chronopoulou"
date: "Lecture 6 Examples"
output: html_document
---


```{r , results='hide'}
bikeshares <-  read.csv("BikeShares.csv", header=TRUE)
dim(bikeshares)
head(bikeshares)
# We remove columns 1,  7, 8, 9, 10:
bikeshares.reg = bikeshares[,c(-1,-7,-8,-9,-10)]
head(bikeshares.reg)
```

<br>
<hr>
<br>

#### Partial F Tests

We start by fitting the Full model, i.e. the model that contains all the model parameters under consideration: `Real Temperature`, `Feels Like Temperature`, `Wind Speed` and `Humidity.`
```{r} 
bikeshare.mlr.full = lm(cnt ~ t1 + t2+ hum + wind_speed, data=bikeshares.reg )
```

Next, we fit the model without the temperature variables (the reduced model):
```{r}
bikeshare.mlr.reduced = lm(cnt ~ hum + wind_speed , data=bikeshares.reg )
```

The partial $F$ test hypothesis is formulated as:
\[
\begin{cases}
&H_0: \beta_1 = \beta_2 = 0\\
&H_{\alpha}: \text{ at least one } \beta \neq 0
\end{cases}
\]

We use tha `anova` function to perform the test:
```{r}
anova(bikeshare.mlr.reduced, bikeshare.mlr.full)
```

We can also perform the test "by hand" as follows:
```{r}
rss.full = sum(bikeshare.mlr.full$res^2)
rss.full
# You can also compute it with
deviance(bikeshare.mlr.full)
rss.reduced = sum(bikeshare.mlr.reduced$res^2)
rss.reduced
deviance(bikeshare.mlr.reduced)
Fstat = (rss.reduced - rss.full)/2/(rss.full/17409)
Fstat
1-pf(Fstat, 2, 17409, lower.tail = FALSE)
```



#### Permutation Tests 


When the normal assumption is questionable, we can use a permutation test.

Suppose our  hypotheses are
\[\begin{cases}
&H_0: bikeshares \sim humidity + wind_speed\\
&H_{\alpha}: bikeshares \sim RealTemp + FeelsLikeTemp + humidity + wind_speed
\end{cases}\]

Let’s use $F$ test statistic from the full model as the test statistic: If `RealTemp` and `FeelsLikeTemp` are indeed relevant to `bikeshares`, we would expect the $F$ test statistic to be large. So a large $F$ test statistic (from the full model) supports $H_{\alpha}$ (i.e., we know the rejection region).

Now the goal is to find the distribution of the $F$ test statistic under the null hypothesis (**without using any normal assumption**). How can we get a data set, which also involves bike shares with all the four variables, and which looks like a data set from $H_0$?

Recall that under $H_0$, `RealTemp` and `FeelsLikeTemp` are irrelevant. In our data, we have 17414 pairs of `(RealTemp, FeelsLikeTemp)`, each corresponding to a Bike Share. What if we randomly assign a pair of `(RealTemp, FeelsLikeTemp)` associated with one bike share to another bike share (repeat this for all 17414 pairs), and then re-fit the full model? The inference based on this new data set should nott be affected that much, if `(RealTemp, FeelsLikeTemp)` is truly irrelevant to `bikeshares`. This is the idea behind the permutation test.


```{r}
n.iter = 2000;
A=numeric(5)
A[1]=2
A
fstats = numeric(n.iter);
class(bikeshares.reg)
head(bikeshares.reg)
for(i in 1:n.iter){
  newbikes = bikeshares.reg;
  newbikes[, c(3,4)] = bikeshares.reg[sample(17414), c(3,4)];
  ge = lm(cnt ~ t1 + t2+ hum + wind_speed, data=newbikes);
  fstats[i] = summary(ge)$fstat[1]
}

# Estimated p-value
length(fstats[fstats > summary(bikeshare.mlr.full)$fstat[1]])/n.iter

## Sample Function
sample(1:5)
sample(1:5)
sample(1:5)
sample(1:5)


```

